The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-08-30 manuscript.Rmd
better writing

Museum collections have been recording our knowledge about species and their geographic ranges, but until recently, this information was fragmented across institutions and difficult to synthesize. The digitization of natural history collections and other biodiversity information systems can now be used to enable synthesis and highlight knowledge gaps. Here we use these data to synthesize the state of our knowledge for marine invertebrates in the continental United States. We show that current knowledge and sampling efforts are inadequate given the diversity of the fauna and the geographic extent of US. While recent high-throughput sequencing technologies promise to greatly speed up the documentation of marine species and their ranges, its value will be greatly improved by anchoring the results into existing DNA sequence data and associated vouchers. However, the best documented phyla have at most 25% of their species sequenced for the most widely used DNA barcoding gene, cytochrome oxidase I (COI). More sampling, coordinated at the scale of the US, is needed to improve our knowledge of species and their geographic ranges. This data is needed to detect the impact of climate changes and other anthropogenic strains to the environment. Advances in computing, information systems, and sequencing technologies, have set the stage to inform where future efforts for documenting biodiversity should be applied.

@@

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood [@Worm2006], and modeling the impact of human activities has been limited to a few taxonomic groups [@Tittensor2010]. For marine organisms, because of the diversity of their complex life history strategies, their size range, and the difficulties associated with sampling their habitats, we have limited information about which species live where, or what portion of their diversity remains to be discovered and described [@Mora2011]. Additionally, coordinated marine ecosystem monitoring initiatives have been lacking [@Duffy2013,@Borja2016] (but see MarineGEO [@Duffy2016marinegeo] and PISCO), making it difficult to monitor changes in species distributions and abundances. While most documented global extinctions of marine species are vertebrates, many others species may have been lost unnoticed because they were not known [@Sala2006]. There is therefore a pressing need to improve our knowledge of marine systems before their functioning are irreversibly altered, and species go extinct.

@@

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), difficult to access, and therefore difficult to synthesize. The development of public databases for specimen records (e.g. iDigBio [@Page2015]), species occurrences (e.g., OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), now allow us to evaluate what and how much we know. Here, we aggregate publicly available data for marine invertebrates that occur within the limits of the exclusive economic zone (EEZ) of the continental United States (US), as well as the Gulf of Mexico (GOM) and the Pacific Northwest (PNW), to highlight gaps in our knowledge that provide a road map for important new directions in global research. We show that while public databases are capturing a significant portion of our knowledge about species richness, most species are known from only a few specimens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but more comprehensive data on species distributions and abundances are needed. These goals can only be achieved with additional cost-effective sampling and a commitment to training more taxonomists and experts that can identify and interpret the material collected.

@@

We extracted `r format(sum(gom_worms$is_marine, na.rm=TRUE), big.mark=",")` and `r format(sum(kozloff_worms$is_marine, na.rm=TRUE), big.mark=",")` marine invertebrate species from species checklists from the Gulf of Mexico [@Tunnell2009] and from the Pacific Northwest [@Kozloff1996] respectively. For the Gulf of Mexico, biodiversity databases included more species for the same geographic area as the one used by the authors of the checklist (iDigBio and OBIS combined: `r format(dplyr::n_distinct(combined_gom_records$worms_valid_name), big.mark=",")` species; iDigBio only: `r format(n_distinct(idigbio_gom_records$worms_valid_name), big.mark=",")` species, $n =$  `r format(nrow(idigbio_gom_records), big.mark=",")` records; OBIS only: `r format(n_distinct(obis_gom_records$worms_valid_name), big.mark=",")` species, $n=$ `r format(nrow(obis_gom_records), big.mark=",")` records). For the PNW, fewer species than included in the regional checklist were recovered from biodiversity databases (iDigBio and OBIS combined: `r format(dplyr::n_distinct(combined_pnw_records$worms_valid_name), big.mark=",")` species; iDigBio only: `r format(dplyr::n_distinct(idigbio_pnw_records$worms_valid_name), big.mark=",")`  species, $n=$ `r format(nrow(idigbio_pnw_records), big.mark=",")` records; OBIS only `r format(dplyr::n_distinct(obis_pnw_records$worms_valid_name), big.mark=",")` species, $n=$ `r format(nrow(obis_pnw_records), big.mark=",")` records). The combination of the species listed in these three sources leads to `r n_distinct(total_gom_records$worms_valid_name)` species for the GOM and `r n_distinct(total_pnw_records$worms_valid_name)` species for the PNW.

@@

Digitization of species richness is higher for the GOM compared with the PNW. A higher proportion of the species included in the regional list for the GOM are captured in biodiversity databases than for the PNW. While `r format_output((3890 + 2663 + 1025 + 316 + 102 + 49 )/n_distinct(total_gom_records$worms_valid_name) * 100)`%  of the species included in the checklist for the GOM have records in at least one of the biodiversity database (Fig. \@ref(fig:database-coverage)A), `r format_output((352 + 292 + 281 + 235 + 76 + 38)/n_distinct(total_pnw_records$worms_valid_name)*100)`% do for the PNW  (Fig. \@ref(fig:database-coverage)B). Additionally, a higher proportion of species occur in the biodiversity databases but are not included in the regional checklist for the GOM compared with the PNW (GOM: `r format_output((1253 + 848 + 556 + 216 + 190 + 124) / n_distinct(total_gom_records$worms_valid_name)*100)`%, $n =$ `r n_distinct(total_gom_records$worms_valid_name)`; PNW:
`r format_output( (281 + 151 + 145 + 80 + 78 + 45) / n_distinct(total_pnw_records$worms_valid_name)*100)`%, $n =$ `r n_distinct(total_pnw_records$worms_valid_name)`).

@@

<!-- need to separate across all phyla statements and by phylum statements -->
We investigated the number of species for each phylum that contributed to the discrepancies between the regional species lists and the database records. We found that the more diverse phyla contributed the most to the differences across the data sources. OBIS tends to be closer to the full record set, but iDigBio has more species for the Gulf of Mexico that are not included in the list (Fig. \@ref(fig:database-overlap)). Many species of mollusks, arthropods, and annelids, known to occur in the Gulf of Mexico or in the Pacific Northwest, have no digital records in these geographic areas.

@@

Several factors may contribute for species to be included in the regional lists but to not have records in biodiversity databases. The first category of biases are linked to property of the species affecting the likelihood of having specimens preserved in museum collections. For instance, rare species may never have been collected or have only been collected a few times, and the corresponding records have not been digitized yet. Species that are difficult to preserve because they are large, very small, or fragile, are deposited in museums less frequently. The other category of biases is associated with data quality for these records. Some of these species may be in the databases but do not have geographic coordinates associated with them, and therefore would not be included in the results of our query.

Missing geographic information contribute to the discrepencies observed.  Using iDigBio we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for the ones that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within the US waters. These numbers indicate that missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Additional digitization of collections, including the capture of their geographic coordinates information, are needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems.

Two factors are most likely to contribute to explain why some species are in databases but were not considered by the authors of the checklists: (1) the species have been documented for the first time since the publication of the checklist, (2) the species records were not accessible or known to the authors.

<!-- include also proportion of species not in list -->
Recently recorded species contribute modestly to this discrepency, as `r sum(recently_collected$gom_in_dbs$min_year > 2001)` species (`r format_output( sum(recently_collected$gom_in_dbs$min_year > 2001)/n_distinct(total_gom_records$worms_valid_name) * 100)`% of the total species richness) have been collected for the first time since the publication of the list in the Gulf of Mexico, and `r sum(recently_collected$pnw_in_dbs$min_year > 1996)` species (`r format_output(sum(recently_collected$pnw_in_dbs$min_year > 1996)/n_distinct(total_pnw_records$worms_valid_name) * 100)`% of the total species richness) for the Pacific Northwest.

The lack to tools for the authors of these checklist to have access to museum records across institutions contributed most to explain why some species were not included in the checklists. These records were difficult to discover before digitization which, for the most part, has occured after the compilation of these checklists in most cases.

@@

<!-- separate idigbio and (and add) obis in the results -->

@@

Further biases exist across phyla and body size. Large-bodied and easily acccessible (e.g. dried specimens like shells, corals, echinoderms) are overrepresented in museum collections, compared with small less accessible organisms, such as the meiofauna. Of `r n_idigbio_phyla` phyla represented in iDigBio, `r records_rare_phyla$n_rare_phyla` have <500 records, and make up `r records_rare_phyla$prop_recs_rare_phyla`% of the iDigBio records. This underrepresentation is also manifested at the species level, while `r records_rare_phyla$prop_spp_rare_phyla_collected`% (n = `r format(records_rare_phyla$total_spp_rare_phyla_collected, big.mark=",")` total species) of the global number of species for these `r records_rare_phyla$n_rare_phyla` phyla occur in iDigBio, this proportion rises to `r records_rare_phyla$prop_spp_common_phyla_collected`% (n = `r format(records_rare_phyla$total_spp_common_phyla_collected, big.mark=",")`) for the `r n_idigbio_phyla - records_rare_phyla$n_rare_phyla` others. Yet, these undersampled taxa, especially those of small body size, are often the most abundant, and are likely to play an important role in the detection of changes in environmental quality[@Bik2012]. Additionally, the ecological implications of changes in the structure and composition of the communities of these small-bodied organisms, on higher trophic levels and ecosystems are largely unexplored.

The proportion of samples that are identified to species in museum collections show a marked decline in recent decades in most taxa other than mollusks, demonstrating limiting taxonomic effort and expertise (Fig. \@ref(fig:identification-through-time)). There is a marked break in the central trend arond the 1970's indicating that it now takes decades, on average, to identify samples deposited in collections. For diverse groups such as annelids or arthropods, the proportion of specimens identified to the species-level is now below 30%.

@@

<!-- add overall mean number before getting into details -->

@@

```{r proportion-barcoded, fig.cap='Proportion of species known to occur in the US EEZ based on iDigBio records, in the Gulf of Mexico [@Tunnell2009], in the Pacific Northwest[@Kozloff1996], that have publicly available DNA barcodes. Numbers next to the bars indicate the number of species for which DNA barcodes are available. (\\#fig:proportion-barcoded)'}
knitr::include_graphics("figures/proportion_barcoded.pdf")
```

## Where do we go from there?

@@

With data aggregation of natural history collections through initiatives like iDigBio, and the compilation of a nearly complete register of marine species with WoRMS, in combination with the development of algorithms for text analysis, it is increasingly easier to validate and spell check the scientific names of the identifications of digitized museum specimens. Linking species names to identifiers in taxonomic databases would facilitate data linking, specimen discoverability, and data usability. Additionally, because species names registers like WoRMS link the history of the names and their synonymies, it is possible to dissociate the concepts of species identification with the concept of species identity, allowing researchers to compare specimens identified at different time periods, a tedious process that would have required expert knowledge in the past.

@@

<!-- separate issues of sampling is needed, and large-scale sampling is best way of doing it --> Coordinated large-scale biodiversity expeditions continue to be needed to improve our understanding of the diversity of species and their spatial distributions. The US marine EEZ is relatively well-known compared to many areas of the world ocean, where studies show that undescribed biodiversity dominates. For example, a recent survey in Western Australia reported that 94.6% of crustaceans and 72% of polychaetes taken at 135 sites were undescribed [@Poore2014].

 However, museum specimens are critical to validate the occurrence of species [@Rocha2014], especially for marine invertebrates for which species limits based on morphology have been known to be challenged by other lines of evidence including molecular data [@Knowlton2000].

@@

Integrating metabarcoding as part of the data collection conducted in biodiversity surveys is a promising approach to screen an habitat for its potential to lead to new discoveries. For instance, by analyzing metabarcoding samples that contain planktonic larvae from a given habitat, we could get an indication of which taxonomic group would benefit most from additional sampling effort or give an indication of the proportion of species that has not been sampled yet. Additionally, data generated by metabarcoding efforts can be re-interpreted as more DNA barcodes associated with correctly identified specimens become available. Building the library of DNA barcodes in parallel of metabarcoding efforts ensures that database matches correspond to actual specimens deposited in natural history collections, ruling out possible artifacts of the method. With such approaches, the geographic distribution of species can build on and be linked to existing records rather than being limited to matching DNA sequences across studies.

@@

For geographic searches, we obtained the boundaries of the Exclusive Economic Zone for the United States from marineregions.org (mrgid:8456) [@mregions]. For the Pacific Northwest, Kozloff did not specify the exact geographic extent covered by his book. We inferred a polygon that encompass part of the Oregon coast to the northern parts of Vancouver island. For the Gulf of Mexico, we used the same geographic area as the one defined in [@Tunnell2009] by modifying manually the eastern boundaries of the shape file for the Gulf of Mexico provided by marineregions.org (mrgid:4288). These geographic areas were split into 0.2° $\times$ 0.2° quadrats that were used to query the web API of iDigBio and OBIS using the R packages ridigbio [@ridigbio-0.3.5] and robis [@robis-0.1.8] respectively. The resulting occurrences were then clipped to only include records that fell within the EEZ boundaries.

>>>>

Museum collections have been recording our knowledge about species and their geographic ranges, but until recently, this information was fragmented across institutions and difficult to synthesize. The digitization of natural history collections and other biodiversity information systems can now be used to enable synthesis and highlight knowledge gaps. Here we use these data to synthesize the state of our knowledge for marine invertebrates in the continental United States. We show that current knowledge and sampling efforts are inadequate given the diversity of the fauna and the geographic extent of US. While recent high-throughput sequencing technologies promise to greatly speed up the documentation of marine species and their ranges, its value will be greatly improved by anchoring the results into existing DNA sequence data and associated vouchers. However, the best documented phyla have at most 25% of their species sequenced for the most widely used DNA barcoding gene, cytochrome oxidase I (COI). More sampling, coordinated at the scale of the US, is needed to improve our knowledge of species and their geographic ranges. This data is needed to detect the impact of climate change and other anthropogenic strains to the environment. Advances in computing, information systems, and sequencing technologies, have set the stage to inform where future efforts for documenting biodiversity should be applied.

@@

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood [@Worm2006], and modeling the impact of human activities has been limited to a few taxonomic groups [@Tittensor2010]. For marine organisms, because of the diversity of their complex life history strategies, their size range, and the difficulties associated with sampling their habitats, we have limited information about which species live where, or what portion of their diversity remains to be discovered and described [@Mora2011]. Additionally, coordinated marine ecosystem monitoring initiatives have been lacking [@Duffy2013,@Borja2016] (but are now getting organized, see MarineGEO [@Duffy2016marinegeo], MBON, and PISCO), making it difficult to monitor changes in species distributions and abundances. While most documented global extinctions of marine species are vertebrates, many others species may have been lost unnoticed because they were not known [@Sala2006]. There is therefore a pressing need to improve our knowledge of marine systems before their functioning are irreversibly altered, and species go extinct.

@@

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), difficult to access, and therefore difficult to synthesize. The development of public databases for specimen records (e.g. iDigBio [@Page2015]), species occurrences (e.g., OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), now allow us to evaluate what and how much we know. Here, we aggregate publicly available data for marine invertebrates that occur within the limits of the exclusive economic zone (EEZ) of the continental United States (US), as well as the Gulf of Mexico (GOM) and the Pacific Northwest (PNW), to highlight gaps in our knowledge that provide a road map for important new directions in global research. We show that while public databases are capturing a significant portion of our knowledge about species richness, most species are known from only a few specimens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on biological communities require not only to get a more accurate list of the species that inhabits the oceans, but also more comprehensive data on species distributions and abundances. These goals can only be achieved with additional cost-effective sampling, and a commitment to training more taxonomists and experts that can identify and interpret the material collected.

@@

We extracted `r format(sum(gom_worms$is_marine, na.rm=TRUE), big.mark=",")` and `r format(sum(kozloff_worms$is_marine, na.rm=TRUE), big.mark=",")` marine invertebrate species from species checklists from the Gulf of Mexico [@Tunnell2009] and from the Pacific Northwest [@Kozloff1996] respectively. For the Gulf of Mexico, biodiversity databases included more species for the same geographic area as the one used by the authors of the checklist (iDigBio and OBIS combined: `r format(dplyr::n_distinct(combined_gom_records$worms_valid_name), big.mark=",")` species; iDigBio only: `r format(n_distinct(idigbio_gom_records$worms_valid_name), big.mark=",")` species, $n =$  `r format(nrow(idigbio_gom_records), big.mark=",")` records; OBIS only: `r format(n_distinct(obis_gom_records$worms_valid_name), big.mark=",")` species, $n=$ `r format(nrow(obis_gom_records), big.mark=",")` records). For the PNW, fewer species than included in the regional checklist were recovered from biodiversity databases (iDigBio and OBIS combined: `r format(dplyr::n_distinct(combined_pnw_records$worms_valid_name), big.mark=",")` species; iDigBio only: `r format(dplyr::n_distinct(idigbio_pnw_records$worms_valid_name), big.mark=",")`  species, $n=$ `r format(nrow(idigbio_pnw_records), big.mark=",")` records; OBIS only `r format(dplyr::n_distinct(obis_pnw_records$worms_valid_name), big.mark=",")` species, $n=$ `r format(nrow(obis_pnw_records), big.mark=",")` records). The combination of the species listed in these three sources leads to `r format(n_distinct(total_gom_records$worms_valid_name), big.mark=",")` species for the GOM and `r format(n_distinct(total_pnw_records$worms_valid_name), big.mark=",")` species for the PNW.

@@

A higher proportion of the species included in the regional list for the GOM are captured in biodiversity databases than for the PNW. While `r format_output((3890 + 2663 + 1025 + 316 + 102 + 49 )/n_distinct(total_gom_records$worms_valid_name) * 100)`%  of the species included in the checklist for the GOM have records in at least one of the biodiversity database (Fig. \@ref(fig:database-coverage)A), `r format_output((352 + 292 + 281 + 235 + 76 + 38)/n_distinct(total_pnw_records$worms_valid_name)*100)`% do for the PNW  (Fig. \@ref(fig:database-coverage)B). Additionally, a higher proportion of species occur in the biodiversity databases but are not included in the regional checklist for the GOM compared with the PNW (GOM: `r format_output((1253 + 848 + 556 + 216 + 190 + 124) / n_distinct(total_gom_records$worms_valid_name)*100)`%, $n =$ `r format(n_distinct(total_gom_records$worms_valid_name), big.mark=",")`; PNW:
`r format_output( (281 + 151 + 145 + 80 + 78 + 45) / n_distinct(total_pnw_records$worms_valid_name)*100)`%, $n =$ `r format(n_distinct(total_pnw_records$worms_valid_name), big.mark=",")`).

@@

The more diverse phyla contribute the most to the discrepancies between the regional species lists and the database records (Fig. \@ref(fig:database-overlap)). OBIS tends to be closer to the full record set, but iDigBio has more species for the Gulf of Mexico that are not included in the list across most phyla. Many species of mollusks, arthropods, and annelids, known to occur in the Gulf of Mexico or in the Pacific Northwest, have no digital records in these geographic areas.

@@

Several factors may contribute for species to be included in the regional lists but to not have records in biodiversity databases. The first category of biases is linked to the likelihood of having specimens for these species in museum collections. For instance, rare species may never have been collected or have only been collected a few times, and the corresponding records have not been digitized yet. Species that are difficult to preserve because they are large, very small, or fragile, are deposited in museums less frequently. The other category of biases is associated with data quality for these records. Some of these species may be in the databases but do not have geographic coordinates associated with them, and therefore would not be included in the results of our query. While it is difficult to assess biases associated with the likelihood of preservation of a species, we investigated how missing geographic coordinates are affecting the discrepancies between the checklists and the databases.

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for the ones that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within the US waters. Additional digitization of collections, including the capture of their geographic coordinates information, are needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems.

Two factors are most likely to contribute to explain why some species are in databases but are not included in the regional the checklists: (1) the species have been documented for the first time since the publication of the checklist, (2) the species records were not accessible or known to the authors.

Recently recorded species contribute modestly to this discrepancy for the Gulf of Mexico, as `r sum(recently_collected$gom_in_dbs$min_year > 2001)` species (`r format_output( sum(recently_collected$gom_in_dbs$min_year > 2001)/n_distinct(total_gom_records$worms_valid_name) * 100)`% of the total species richness, `r format_output(sum(recently_collected$gom_in_dbs$min_year > 2001)/spp_not_in_lists$gom_not_in_list * 100)`% of the species not on the checklist),  have been collected for the first time since the publication of the checklist, but more significantly for the Pacific Northwest with  `r sum(recently_collected$pnw_in_dbs$min_year > 1996)` species (`r format_output(sum(recently_collected$pnw_in_dbs$min_year > 1996)/n_distinct(total_pnw_records$worms_valid_name) * 100)`% of the total species richness,  `r format_output(sum(recently_collected$pnw_in_dbs$min_year > 1996)/spp_not_in_lists$pnw_not_in_list * 100)`% of the species not on the checklist).

The lack of tools for the authors of these checklist to have access to museum records across institutions contributes most to explain why some species were not included in the checklists. Many of these records were difficult to discover before digitization which, for the most part, has occurred after the compilation of these checklists. We estimate that using biodiversity databases increases the number of species included in the checklist by `r format_output(spp_not_in_lists$gom_not_in_list/n_distinct(total_gom_species$worms_valid_name)*100)`% for the Gulf of Mexico, and `r format_output(spp_not_in_lists$pnw_not_in_list/n_distinct(total_pnw_species$worms_valid_name)*100)`% for the Pacific Northwest.

@@

@@

Further biases exist across phyla and body size. Large-bodied and easily accessible (e.g. dried specimens like shells, corals, echinoderms) are over-represented in museum collections, compared with small less accessible organisms, such as the meiofauna. Of `r n_idigbio_phyla` phyla represented in iDigBio, `r records_rare_phyla$n_rare_phyla` have <500 records, and make up `r records_rare_phyla$prop_recs_rare_phyla`% of the iDigBio records. This under-representation is also manifested at the species level, while `r records_rare_phyla$prop_spp_rare_phyla_collected`% (n = `r format(records_rare_phyla$total_spp_rare_phyla_collected, big.mark=",")` total species) of the global number of species for these `r records_rare_phyla$n_rare_phyla` phyla occur in iDigBio, this proportion rises to `r records_rare_phyla$prop_spp_common_phyla_collected`% (n = `r format(records_rare_phyla$total_spp_common_phyla_collected, big.mark=",")`) for the `r n_idigbio_phyla - records_rare_phyla$n_rare_phyla` others. Yet, these undersampled taxa, especially those of small body size, are often the most abundant, and are likely to play an important role in the detection of changes in environmental quality[@Bik2012]. Additionally, the ecological implications of changes in the structure and composition of the communities of these small-bodied organisms, on higher trophic levels and ecosystems are largely unexplored.

The proportion of samples that are identified to species in museum collections show a marked decline in recent decades in most taxa other than mollusks, demonstrating limiting taxonomic effort and expertise (Fig. \@ref(fig:identification-through-time)). There is a marked break in the central trend around the 1970's indicating that it now takes decades, on average, to identify samples deposited in collections. For diverse groups such as annelids or arthropods, the proportion of specimens identified to the species-level is now below 30%.

@@

@@

```{r proportion-barcoded, fig.cap='Proportion of species with publicly available DNA barcodes  known to occur the US EEZ (iDigBio and OBIS records), in the Gulf of Mexico (iDigBio, OBIS, and checklist[@Tunnell2009]), and in the Pacific Northwest (iDigBio, OBIS, and checklist[@Kozloff1996]). Numbers next to the bars indicate the number of species for which DNA barcodes are available. (\\#fig:proportion-barcoded)'}
knitr::include_graphics("figures/proportion_barcoded.pdf")
```

## Discussion

@@

With data aggregation of natural history collections through initiatives like iDigBio, and the compilation of a nearly complete register of marine species with WoRMS, in combination with the development of algorithms for text analysis, it is increasingly easier to validate and spell check the scientific names of the identifications of digitized museum specimens. Linking species names to identifiers in taxonomic databases would facilitate data linking, specimen discoverability, and data usability. Additionally, because species names registers like WoRMS link the history of the names and their synonymies, it is possible to dissociate the concepts of species identification with the concept of species identity, allowing researchers to compare specimens identified at different time periods.

@@

<!-- separate issues of sampling is needed, and large-scale sampling is best way of doing it -->

<!-- Coordinated large-scale biodiversity expeditions continue to be needed to improve our understanding of the diversity of species and their spatial distributions. The US marine EEZ is relatively well-known compared to many areas of the world ocean, where studies show that undescribed biodiversity dominates.  -->

<!-- For example, a recent survey in Western Australia reported that 94.6% of crustaceans and 72% of polychaetes taken at 135 sites were undescribed [@Poore2014]. -->

Museum specimens are critical to validate the occurrence of species [@Rocha2014], especially for marine invertebrates for which species limits based on morphology have been known to be challenged by other lines of evidence including molecular data [@Knowlton2000].

@@

Integrating metabarcoding as part of the data collection conducted in biodiversity surveys is a promising approach to screen an habitat for its potential to lead to new discoveries. For instance, by analyzing metabarcoding samples that contain planktonic larvae from a given habitat, we could get an indication of which taxonomic group would benefit most from additional sampling effort or give an indication of the proportion of species that has not been sampled yet. Additionally, data generated by metabarcoding efforts can be re-interpreted as more DNA barcodes associated with correctly identified specimens become available. Building the library of DNA barcodes in parallel of metabarcoding efforts ensures that database matches correspond to actual specimens deposited in natural history collections, ruling out possible artifacts of the method. With such approaches, the geographic distribution of species can build on and be linked to existing records, species and their natural history, rather than being limited to matching DNA sequences across studies.

@@

For geographic searches in OBIS and iDigBio, we obtained the boundaries of the Exclusive Economic Zone for the United States from marineregions.org (mrgid:8456) [@mregions]. For the Pacific Northwest, Kozloff did not specify the exact geographic extent covered by his book. We inferred a polygon that encompass part of the Oregon coast to the northern parts of Vancouver island. For the Gulf of Mexico, we used the same geographic area as the one defined in [@Tunnell2009] by modifying manually the eastern boundaries of the shape file for the Gulf of Mexico provided by marineregions.org (mrgid:4288). These geographic areas were split into 0.2° $\times$ 0.2° quadrats that were used to query the web API of iDigBio and OBIS using the R packages ridigbio [@ridigbio-0.3.5] and robis [@robis-0.1.8] respectively. The resulting occurrences were then clipped to only include records that fell within the EEZ boundaries.
